---
title: "Activity Monitoring Project"
author: "Maya Acharya"
date: "November 10, 2014"
output: pdf_document
---
#Loading and Processing Data
Use read.csv to load the data into a variable, called **activity**.
Print the column names of the data to see what you're working with.
```{r load_data, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(chron)
library(lattice)
library(ggplot2)
activity <- read.csv("activity.csv")
activity <- tbl_df(activity)
colnames(activity)
```
#Summarizing the steps data
Make a histogram of the *steps* column of **activity** to visualize the data.
```{r step_hist}
hist(activity$steps, xlab="Steps")
```
  
###Find the *mean* and *median* of the *steps* data.
Use the *mean* and *median* functions on the **steps** column of the **activity** data.
```{r step_mm}
step_mean <- mean(activity$steps, na.rm=TRUE)
step_med <- median(activity$steps, na.rm=TRUE)
{print(step_mean)
 print(step_med)}
```
###Mean of *steps* data: `r step_mean`  
###Median of *steps* data: `r step_med`  

#Visualizing time intervals over days
Group the data by interval. Use the *summarize* function to group the **activity** data, and find the mean of each interval.
```{r time_intervals}
time_interval_means <- summarize(group_by(activity, interval), mean(steps, na.rm=TRUE))
```

  
  Use the *base plotting* system to plot the average steps by time interval.
```{r plot_time_intervals}
plot(time_interval_means, type="l",xlab="Time Interval", ylab = "Average Steps")
```

  
To find the time interval with the highest average steps, use the *arrange* function to arrange the **time_interval_means** by descending average steps. The first row will be the time interval with the highest average steps. 
```{r greatest_interval, message=FALSE, warning=FALSE}
colnames(time_interval_means) <- c("interval", "steps")
time_interval_means <- arrange(time_interval_means, desc(steps))
```
  
###Interval with highest average steps: `r time_interval_means[1,1]` 
(`r time_interval_means[1,2]` steps)

#Imputing Missing Values
To impute missing values, I made each NA steps value equal to the average steps value for that time interval.
```{r impute_missing_values}
time_interval_means <- arrange(time_interval_means, interval)
activity_imputed <- activity
for(i in 1:nrow(activity)){
  if(is.na(activity[i,1])){
    activity_imputed[i,1] <- as.numeric(time_interval_means[(activity[i,3]/5)+1,2])
  }
}
hist(activity_imputed$steps, xlab="Steps")
step_mean_imputed <- mean(activity_imputed$steps, na.rm=TRUE)
step_med_imputed <- median(activity_imputed$steps, na.rm=TRUE)
{print(step_mean_imputed)
 print(step_med_imputed)}
```
###Mean of *steps* data with imputed missing values: `r step_mean_imputed`  
###Median of *steps* data with imputed missing values: `r step_med_imputed` 

#Looking at Weekdays vs. Weekends
Use *dplyr*'s mutate function to split the **date** column into two factors, **TRUE** for days that are weekends, or **FALSE** for days that are during the work week. Store this in a new column called **day**. Then, use the *factor* function to relabel **day** with **Weekday** and **Weekend**. Use *lattice*'s *xyplot* function to create a time plot of the average steps per interval split over **Weekend**s and **Weekday**s
```{r weekdays}
dayActivity <- mutate(activity, day=is.weekend(as.Date(activity$date)))
dayActivity$day <- factor(dayActivity$day, label=c("Weekday", "Weekend"))
dayActivity <- summarize(group_by(dayActivity, interval, day), mean(steps, na.rm=TRUE))
colnames(dayActivity) <- c("interval", "day", "steps")
xyplot(steps ~ interval | day, col="#68ab98", type="l", layout=c(1,2), data=dayActivity)
```
  
    
There is a clearly visible difference between the **steps** data for weekends and for weekdays. On weekends, the steps taken between the $500^{th}$ and $2000^{th}$ time intervals are consistently higher than the other time intervals. On weekdays, there is a sizeable spike from the $500^{th}$ to the $1000^{th}$ time interval.
